---
title: "Fundamentals of Computing and Data Display"
subtitle: "Exercise"
author: "Christoph Kern"
output: html_notebook
Student: Abigail Kamp
---

## Setup

```{r}
library(xml2)
library(rvest)
```

## Web Scraping

In this notebook, your task is to scrape some information from Wikipedia. We start with the following page about Grand Boulevard, a Chicago Community Area.

https://en.wikipedia.org/wiki/Grand_Boulevard,_Chicago

The ultimate goal is to gather the table "Historical population" and convert it to a `data.frame`.

As a first step, read in the html page as an R object.

```{r}
url <- read_html("https://en.wikipedia.org/wiki/Grand_Boulevard,_Chicago")
```

Now extract the tables from this object (using the `rvest` package) and save the result as a new object. Follow the instructions if there is an error.

```{r}
tables <- html_nodes(url, "table")
html_table(tables, fill = TRUE)
```

Use `str()` on this new object -- it should be a list. Try to find the position of the "Historical population" in this list, since we need it in the next step.

```{r}
str(tables)
```

Extract the "Historical population" table from the list and save it as another object. You can use subsetting via `[[...]]` to extract pieces from a list. Print the result.

```{r}
pop<-html_table(tables[2],fill=TRUE, trim=TRUE)
pop
```

You will see that the table needs some additional formatting. We only want rows and columns with actual values (I called the table object `pop`).

```{r}
pop <- pop[[1]]
pop <-pop[2:10, -3]
pop
```

That's it for this page. However, we may want to repeat this process for other community areas. The Wikipedia page https://en.wikipedia.org/wiki/Grand_Boulevard,_Chicago has a section on "Places adjacent to Grand Boulevard, Chicago" at the bottom. Can you find the corresponding table in the list of tables that you created earlier? Extract this table as a new object.
    
```{r}
all_tables <- url %>%
  html_nodes("table") %>%
  html_table(fill=TRUE)

all_tables <- all_tables[[3]]
all_tables

```

Now, grab the community areas east of Grand Boulevard and save them as a character vector. Print the result.

```{r}
east <- all_tables[-1,3]
print(east)
```

We want to use this list to create a loop that extracts the population tables from the Wikipedia pages of these places. To make this work and build valid urls, we need to replace empty spaces in the character vector with underscores. This can be done with `gsub()`, or by hand.

```{r}
east <- gsub(" ", "_", east)
east
```

The resulting vector should look like this:  
"Oakland,_Chicago" "Kenwood,_Chicago" "Hyde_Park,_Chicago"

To prepare the loop, we also want to copy our `pop` table and rename it as `pops`. In the loop we append this table by adding columns from the other community areas.

```{r}
pops <- pop
names(pops) <- c("Year", "Pop", "%+-")

```

Build a small loop to test whether you can build valid urls using the vector of places and pasting each element of it after `https://en.wikipedia.org/wiki/` in a for loop. Calling `url` shows the last url of this loop, which should be https://en.wikipedia.org/wiki/Hyde_Park,_Chicago.

```{r}
#I created a loop for each url, but I can't get rid of the extra text. It kind of screws with my results but I can't figure out how to get rid of it
url <- vector("character")

for(i in east) {
  url[i] <- paste("https://en.wikipedia.org/wiki/", i, sep = "")
}
url
```

Now, extend the loop and add the code that is needed to grab the demographic tables from each page. Add columns to the original table `pops` using `cbind()`.

-- If you don't know the exact R functions or how to put them together, it is also ok to write pseudo-code. --

```{r}

total <- 
  lapply(url, function(i) {
  webpage <- read_html(i)
  city <-html_nodes(webpage, 'table')
  city_pop <- html_table(city, fill=TRUE, trim = TRUE)[[2]]
  city_pop[3:11,]
  city_pop[,3] <- NULL
  names(city_pop) <- c("Year", "Pop", "%+-")
  as.data.frame(city_pop)
  return(city_pop)
})

total

names(total) <- c(east)

dim(pop)
dim(total)

#I know that I could bind them with cbind, but it throws an error becuase I have the extra names and couldn't figure out how to get rid of them
#to allow the lower code cunks to run, I've commented out the line below
#cbind(pops, total)
```

Use `str()` on the final result.

```{r}
str(total)
```

And print it.

```{r}
print(total)

```
